---
title: "Performance Metrics"
metadata:
  robots: noindex,nofollow
format: 
  html:
    self-contained: true   # embed all images, CSS, JS in the HTML
    include-in-header:
      text: |
        <link rel="icon" href="favicon.jpg">
    toc: true
    toc-depth: 4
    toc-location: right
---

<link href="https://fonts.googleapis.com/css2?family=Roboto:wght@400;500;700;900&display=swap" rel="stylesheet">

<style>

body {
  font-family: "Roboto", Arial, sans-serif;
  font-size: 16px;
  color: #333;
}

h1, h2, h3, h4, h5, h6 {
  font-family: "Roboto", Arial, sans-serif;
  font-weight: 500;
}

h1, h2, h3 {
  font-weight: 900;
}

img {
  width: 100%;
}

.dashboard-footer {
  margin-top: 2rem;
  padding-top: 1.5rem;
  border-top: 1px solid #ddd;
  font-size: 0.85rem;
  color: #666;
  text-align: center;
}

</style>


```{r, echo=FALSE, message=FALSE, warning=FALSE}

here::i_am("performance_metrics.qmd")

```

## Ridership  

Ridership is a key indicator of transit demand and usage.

<iframe title="Total Ridership" aria-label="Line chart" id="datawrapper-chart-WlAlx" src="https://datawrapper.dwcdn.net/WlAlx/1/" scrolling="no" frameborder="0" style="width: 0; min-width: 100% !important; border: none;" height="439" data-external="1"></iframe><script type="text/javascript">(function(){function e(){window.addEventListener(`message`,function(e){if(e.data[`datawrapper-height`]!==void 0){var t=document.querySelectorAll(`iframe`);for(var n in e.data[`datawrapper-height`])for(var r=0,i;i=t[r];r++)if(i.contentWindow===e.source){var a=e.data[`datawrapper-height`][n]+`px`;i.style.height=a}}})}e()})();</script>

Ridership data by route can be found in the [ridership tracker](https://greenmountaintransit.github.io/opendata/site/).  

## Productivity  

Ridership per vehicle revenue hour measures how effectively service hours are being utilized by riders.

<iframe title="Ridership per Vehicle Revenue Hour" aria-label="Line chart" id="datawrapper-chart-Gz99h" src="https://datawrapper.dwcdn.net/Gz99h/1/" scrolling="no" frameborder="0" style="width: 0; min-width: 100% !important; border: none;" height="439" data-external="1"></iframe><script type="text/javascript">(function(){function e(){window.addEventListener(`message`,function(e){if(e.data[`datawrapper-height`]!==void 0){var t=document.querySelectorAll(`iframe`);for(var n in e.data[`datawrapper-height`])for(var r=0,i;i=t[r];r++)if(i.contentWindow===e.source){var a=e.data[`datawrapper-height`][n]+`px`;i.style.height=a}}})}e()})();</script>  

Ridership per vehicle revenue hour data by route can be found in the [ridership tracker](https://greenmountaintransit.github.io/opendata/site/).  

## Reliability  

On-time performance reflects the consistency and dependability of our service. 
On-time is defined as an arrival occurring no more than 1 minute early and no 
more than 5 minutes late relative to the scheduled arrival time.

<iframe title="On-time Performance" aria-label="Line chart" id="datawrapper-chart-A12sD" src="https://datawrapper.dwcdn.net/A12sD/2/" scrolling="no" frameborder="0" style="width: 0; min-width: 100% !important; border: none;" height="454" data-external="1"></iframe><script type="text/javascript">(function(){function e(){window.addEventListener(`message`,function(e){if(e.data[`datawrapper-height`]!==void 0){var t=document.querySelectorAll(`iframe`);for(var n in e.data[`datawrapper-height`])for(var r=0,i;i=t[r];r++)if(i.contentWindow===e.source){var a=e.data[`datawrapper-height`][n]+`px`;i.style.height=a}}})}e()})();</script>

## Data quality  

Data completeness measures the accuracy and reliability of data used for real-time passenger information and service planning, while also demonstrating the utilization and value of GMT's transit technology investments.

<iframe title="Data Completeness" aria-label="Stacked column chart" id="datawrapper-chart-ZEdGM" src="https://datawrapper.dwcdn.net/ZEdGM/1/" scrolling="no" frameborder="0" style="width: 0; min-width: 100% !important; border: none;" height="458" data-external="1"></iframe><script type="text/javascript">(function(){function e(){window.addEventListener(`message`,function(e){if(e.data[`datawrapper-height`]!==void 0){var t=document.querySelectorAll(`iframe`);for(var n in e.data[`datawrapper-height`])for(var r=0,i;i=t[r];r++)if(i.contentWindow===e.source){var a=e.data[`datawrapper-height`][n]+`px`;i.style.height=a}}})}e()})();</script>

## Missed Service 

Dropped runs represent scheduled trips that could not be operated as planned.

<iframe title="Dropped Runs" aria-label="Stacked column chart" id="datawrapper-chart-uLNyj" src="https://datawrapper.dwcdn.net/uLNyj/1/" scrolling="no" frameborder="0" style="width: 0; min-width: 100% !important; border: none;" height="492" data-external="1"></iframe><script type="text/javascript">(function(){function e(){window.addEventListener(`message`,function(e){if(e.data[`datawrapper-height`]!==void 0){var t=document.querySelectorAll(`iframe`);for(var n in e.data[`datawrapper-height`])for(var r=0,i;i=t[r];r++)if(i.contentWindow===e.source){var a=e.data[`datawrapper-height`][n]+`px`;i.style.height=a}}})}e()})();</script>

## Customer Feedback  

```{r, echo=FALSE, message=FALSE, warning=FALSE}

cs_per_vh_per_month <- readr::read_csv(here::here("data/customer_service/customer_service_interactions.csv"))

fy_to_date_interactions_per_1000_vrh <- scales::comma(
  (sum(cs_per_vh_per_month$cs_interactions) / sum(cs_per_vh_per_month$total_vrh)) * 1000,
  accuracy = .1
)

```

Customer service contacts provide insight into rider questions, concerns, and feedback. Fiscal year-to-date, 
GMT has recorded an average of **`r fy_to_date_interactions_per_1000_vrh`** customer service interactions for every 1,000 hours of service operated.

<iframe title="Customer Service Contacts" aria-label="Stacked column chart" id="datawrapper-chart-tDQ69" src="https://datawrapper.dwcdn.net/tDQ69/1/" scrolling="no" frameborder="0" style="width: 0; min-width: 100% !important; border: none;" height="509" data-external="1"></iframe><script type="text/javascript">(function(){function e(){window.addEventListener(`message`,function(e){if(e.data[`datawrapper-height`]!==void 0){var t=document.querySelectorAll(`iframe`);for(var n in e.data[`datawrapper-height`])for(var r=0,i;i=t[r];r++)if(i.contentWindow===e.source){var a=e.data[`datawrapper-height`][n]+`px`;i.style.height=a}}})}e()})();</script>


<div class="dashboard-footer">
  © 2026 [Green Mountain Transit](https://ridegmt.com/). Data provided as-is and subject to revision.
</div>


