---
title: "GMT Ridership Tracker"
format: 
  html:
    self-contained: true   # embed all images, CSS, JS in the HTML
    include-in-header:
      text: |
        <link rel="icon" href="favicon.jpg">
    toc: true
    toc-depth: 4
    toc-location: right
---

<link href="https://fonts.googleapis.com/css2?family=Roboto:wght@400;500;700;900&display=swap" rel="stylesheet">

<style>

body {
  font-family: "Roboto", Arial, sans-serif;
  font-size: 16px;
  color: #333;
}

h1, h2, h3, h4, h5, h6 {
  font-family: "Roboto", Arial, sans-serif;
  font-weight: 500;
}

h1, h2, h3 {
  font-weight: 900;
}

img {
  width: 100%;
}

.dashboard-footer {
  margin-top: 2rem;
  padding-top: 1.5rem;
  border-top: 1px solid #ddd;
  font-size: 0.85rem;
  color: #666;
  text-align: center;
}

</style>

```{r, echo=FALSE, message=FALSE, warning=FALSE}

library(dplyr)
library(tidyr)
library(here)
library(ggplot2)
library(gmtviz)
library(DT)
library(stringr)
library(readr)
library(googledrive)
library(DatawRappr)

load_fonts()

i_am("ridership_tracker.qmd")

```


```{r, message=FALSE, echo=FALSE, warning=FALSE}

publish_to_dw <- TRUE
sync_with_gdrive <- TRUE

if (sync_with_gdrive) {
  drive_auth()
}

```


```{r, message=FALSE, echo=FALSE}

fy26_urban_sirs_stats_fpath <- here("data/ridership/fy26_urban_ridership.csv")
fy25_urban_sirs_stats_fpath <- here("data/ridership/fy25_urban_ridership.csv")

if (sync_with_gdrive) {
  
  fy26_urban_gdrive <- drive_get(paste0("coredata/processed/fy26_urban_ridership.csv"))
  fy25_urban_gdrive <- drive_get(paste0("coredata/processed/fy25_urban_ridership.csv"))
  
  drive_download(fy26_urban_gdrive, path = fy26_urban_sirs_stats_fpath, overwrite = TRUE)
  drive_download(fy25_urban_gdrive, path = fy25_urban_sirs_stats_fpath, overwrite = TRUE)
  
}

fy25_urban_sirs_stats <- read_csv(fy25_urban_sirs_stats_fpath) %>%
  mutate(route = as.character(route))

fy26_urban_sirs_stats <- read_csv(fy26_urban_sirs_stats_fpath) %>%
  mutate(route = as.character(route))

# Combine them together
historic_gmt_ridership <- bind_rows(
  fy25_urban_sirs_stats %>% mutate(fiscal_year = 2025),
  fy26_urban_sirs_stats %>% mutate(fiscal_year = 2026)
)

```


```{r, echo=FALSE, message=FALSE}

month_order <- c("Jul", "Aug", "Sep", "Oct", "Nov", "Dec",
                 "Jan", "Feb", "Mar", "Apr", "May", "Jun")

chrt_dta <- historic_gmt_ridership %>%
  mutate(
    route = if_else(
      route %in% c("56", "96"),
      "56-96",
      route
    )
  ) %>%
  # Group to consolidate 56 and 96
  group_by(fiscal_year, month, route) %>%
  summarise(
    across(matches("(_vrh|_ridership)$"), ~ sum(.x, na.rm = TRUE)),
    # across(all_of(c("weekdays", "saturdays", "sundays", "total_service_days")), ~ max(.x, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(
    month = substr(month, 1, 3), # keep first 3 letters
    month = factor(month, levels = month_order),
    fiscal_year = as.character(fiscal_year),
    ridership_per_vrh = total_ridership / total_vrh
  )

```


```{r, message=FALSE, echo=FALSE}

get_data_table <- function(chrt_dta, route) {
  
  chrt_dta %>%
    mutate(
      month = factor(month, levels = month_order)
    ) %>%
    select(route, month, fiscal_year, total_ridership, total_vrh, ridership_per_vrh) %>%
    rename(
      Route = route,
      `Fiscal Year` = fiscal_year,
      Month = month,
      `Total Ridership` = total_ridership,
      `Total Vehicle Revenue Hours` = total_vrh,
      `Ridership per VRH` = ridership_per_vrh
    ) %>%
    filter(Route == route) %>%
    arrange(desc(`Fiscal Year`), desc(Month)) %>%
    datatable(rownames = FALSE, options = list(pageLength = 5)) %>%
    formatRound("Total Vehicle Revenue Hours", 1) %>%
    formatRound("Ridership per VRH", 1) %>%
    formatRound("Total Ridership", 1)
  
}

```


```{r, message=FALSE, echo=FALSE}

get_line_chart <- function(dta, y_var, color_palette) {
  
  fig <- dta %>%
    ggplot(
      aes(x = month, y = !!as.symbol(y_var), group = fiscal_year)
    ) +
    geom_line(aes(color = fiscal_year), linewidth = 1.5) +
    geom_point(aes(color = fiscal_year), size = 3) +
    scale_color_brewer(palette = color_palette) +
    scale_y_continuous(
      expand = expansion(mult = c(0, .1)),
      limits = c(0,NA),
      labels = scales::label_comma(accuracy = 1)
    ) +
    scale_x_discrete(
      expand = expansion(mult = c(0.01, 0.01)),
      limits = month_order
    ) +
    theme_gmt(base_size = 10) +
    theme(
      axis.text.x = element_text(color = "dimgray"),
      axis.text.y = element_text(color = "dimgray"),
      axis.ticks = element_line(linewidth = .25, color = "#d0d2ce"),
      plot.margin = margin(20, 20, 20, 20, "pt")
    ) +
    labs(
      title = NULL,
      subtitle = NULL,
      x = NULL,
      y = NULL
    )
  
  return(fig)
  
}

```


```{r, echo=FALSE, message=FALSE}

totals_dta <- chrt_dta %>%
  group_by(fiscal_year, month) %>%
  summarise(
    total_ridership = sum(total_ridership),
    total_vrh = sum(total_vrh)
  ) %>%
  ungroup() %>%
  mutate(
    month = factor(month, levels = month_order),
    ridership_per_vrh = total_ridership / total_vrh
  )

```


```{r, echo=FALSE, message=FALSE}

# Total ridership
fig <- get_line_chart(
  totals_dta,
  "total_ridership",
  1
)
  
save_plot(fig, here(paste0("export/gmt_ridership_over_time_all_routes.svg")), chart_height = 2.5)

```


```{r, echo=FALSE, message=FALSE}

# Vehicle Revenue Hour
fig <- get_line_chart(
  totals_dta,
  "total_vrh",
  7
)
  
save_plot(fig, here(paste0("export/gmt_vrh_over_time_all_routes.svg")), chart_height = 2.5)

```


```{r, echo=FALSE, message=FALSE}

# Ridership per vehicle revenue hour
fig <- get_line_chart(
  totals_dta,
  "ridership_per_vrh",
  2
)

save_plot(fig, here(paste0("export/gmt_ridership_per_vrh_over_time_all_routes.svg")), chart_height = 2.5)

```


```{r, echo=FALSE, warning=FALSE, message=FALSE}

## Datawrapper charts
dw_dta <- totals_dta %>%
  mutate(
    month_num = match(month, month.abb),
    date = as.Date(
      ifelse(
        month_num >= 7,
        paste0("2000-", sprintf("%02d", match(month, month.abb)), "-01"),
        paste0("2001-", sprintf("%02d", match(month, month.abb)), "-01")
      )
    )
  )

ridership_dw <- dw_dta %>%
  select(date, total_ridership, fiscal_year) %>%
  pivot_wider(
    names_from = "date",
    values_from = "total_ridership"
  )

# write.csv(ridership_dw, stdout(), row.names = FALSE)

dw_chart_id <- "WlAlx"
if (publish_to_dw) {
  dw_data_to_chart(ridership_dw, dw_chart_id, parse_dates = FALSE)
  dw_publish_chart(dw_chart_id, return_urls = FALSE) 
}


ridership_per_vrh_dw <- dw_dta %>%
  select(date, ridership_per_vrh, fiscal_year) %>%
  pivot_wider(
    names_from = "date",
    values_from = "ridership_per_vrh"
  )

# write.csv(ridership_per_vrh_dw, stdout(), row.names = FALSE)

dw_chart_id <- "Gz99h"
if (publish_to_dw) {
  dw_data_to_chart(ridership_per_vrh_dw, dw_chart_id, parse_dates = FALSE)
  dw_publish_chart(dw_chart_id, return_urls = FALSE) 
}

```


```{r, echo=FALSE, warning=FALSE}

routes <- chrt_dta %>% pull(route) %>% unique()
months <- chrt_dta %>% pull(month) %>% unique()

for (gmt_route in routes) {
  
  ## Total Ridership
  fig <- get_line_chart(
    chrt_dta %>% filter(route == gmt_route),
    "total_ridership",
    1
  )
  
  save_plot(fig, here(paste0("export/gmt_ridership_over_time_route_", gmt_route, ".svg")), chart_height = 2.5)
  
  
  ## Vehicle Revenue Hours
  fig <- get_line_chart(
    chrt_dta %>% filter(route == gmt_route),
    "total_vrh",
    7
  )
  
  save_plot(fig, here(paste0("export/gmt_vrh_over_time_route_", gmt_route, ".svg")), chart_height = 2.5)  
  
  ## Ridership per Vehicle Revenue Hour
  fig <- get_line_chart(
    chrt_dta %>% filter(route == gmt_route),
    "ridership_per_vrh",
    2
  )
  
  save_plot(fig, here(paste0("export/gmt_ridership_per_vrh_over_time_route_", gmt_route, ".svg")), chart_height = 2.5)
  
}

```

This dashboard displays ridership trends for [Green Mountain Transit](https://ridegmt.com/)’s urban service area (Chittenden County).

Three metrics are presented for each route, month, and fiscal year:  

- **Total ridership**: The total number of recorded passenger boardings  
- **Total vehicle revenue hours**: The total time that vehicles were in service    
- **Ridership per vehicle revenue hour**: A productivity measure calculated as total ridership divided by total vehicle revenue hours  

GMT's fiscal year begins July 1.

Data downloads are available at <a href="index.html" >GMT’s open data site</a>.

If you have feedback or questions, please complete our [feedback form](https://docs.google.com/forms/d/e/1FAIpQLSdzSlkfUjXiXvld0SWsigx7xpIfNXpNgof8Y3_VfITtoDC8nw/viewform) or contact **planning@ridegmt.com**.

---  


### All Routes

<em>Note: All Route totals do not include Shopping Special or School Tripper ridership.</em>  

#### Total ridership 

![](export/gmt_ridership_over_time_all_routes.svg){fig-alt="Line chart of total GMT ridership over time"}  

#### Total vehicle revenue hours 

![](export/gmt_vrh_over_time_all_routes.svg){fig-alt="Line chart of total GMT vehicle revenue hours over time"}  

#### Ridership per vehicle revenue hour 

![](export/gmt_ridership_per_vrh_over_time_all_routes.svg){fig-alt="Line chart of GMT ridership per vehicle revenue hour over time"}  

#### Detailed data  

```{r, echo=FALSE}

# For All Routes
totals_dta %>%
  mutate(
    month = factor(month, levels = month_order)
  ) %>%
  rename(
    `Ridership per VRH` = ridership_per_vrh,
    `Fiscal Year` = fiscal_year,
    `Total Vehicle Revenue Hours` = total_vrh,
    `Total Ridership` = total_ridership,
    `Month` = month
  ) %>%
  arrange(desc(`Fiscal Year`), desc(Month)) %>%
  datatable(rownames = FALSE, options = list(pageLength = 5)) %>%
  formatRound("Total Vehicle Revenue Hours", 1) %>%
  formatRound("Ridership per VRH", 1) %>%
  formatRound("Total Ridership", 1)

```

---  

### 1: Williston

#### Total ridership 

![](export/gmt_ridership_over_time_route_1.svg){fig-alt="Line chart of GMT route 1 ridership over time"}

#### Total vehicle revenue hours

![](export/gmt_vrh_over_time_route_1.svg){fig-alt="Line chart of GMT route 1 vehicle revenue hours over time"}

#### Ridership per vehicle revenue hour 

![](export/gmt_ridership_per_vrh_over_time_route_1.svg){fig-alt="Line chart of GMT route 1 ridership per vehicle revenue hour over time"}

#### Detailed data  

```{r, echo=FALSE}
get_data_table(chrt_dta, "1")
```

---  


### 2: Essex Junction    

#### Total ridership 

![](export/gmt_ridership_over_time_route_2.svg){fig-alt="Line chart of GMT route 2 ridership over time"}

#### Total vehicle revenue hours

![](export/gmt_vrh_over_time_route_2.svg){fig-alt="Line chart of GMT route 2 vehicle revenue hours over time"}

#### Ridership per vehicle revenue hour 

![](export/gmt_ridership_per_vrh_over_time_route_2.svg){fig-alt="Line chart of GMT route 2 ridership per vehicle revenue hour over time"}

#### Detailed data  

```{r, echo=FALSE}
get_data_table(chrt_dta, "2")
```

---  


### 4: Essex Center    

Note: Route 4 Essex Center was introduced in June 2025.

#### Total ridership 

![](export/gmt_ridership_over_time_route_4.svg){fig-alt="Line chart of GMT route 4 ridership over time"}

#### Total vehicle revenue hours

![](export/gmt_vrh_over_time_route_4.svg){fig-alt="Line chart of GMT route 4 vehicle revenue hours over time"}

#### Ridership per vehicle revenue hour 

![](export/gmt_ridership_per_vrh_over_time_route_4.svg){fig-alt="Line chart of GMT route 4 ridership per vehicle revenue hour over time"}

#### Detailed data  

```{r, echo=FALSE}
get_data_table(chrt_dta, "4")
```

---  


### 5: Pine Street  

#### Total ridership 

![](export/gmt_ridership_over_time_route_5.svg){fig-alt="Line chart of GMT route 5 ridership over time"}

#### Total vehicle revenue hours

![](export/gmt_vrh_over_time_route_5.svg){fig-alt="Line chart of GMT route 5 vehicle revenue hours over time"}

#### Ridership per vehicle revenue hour 

![](export/gmt_ridership_per_vrh_over_time_route_5.svg){fig-alt="Line chart of GMT route 5 ridership per vehicle revenue hour over time"}

#### Detailed data  

```{r, echo=FALSE}
get_data_table(chrt_dta, "5")
```


---  


### 6: Shelburne Road  

#### Total ridership 

![](export/gmt_ridership_over_time_route_6.svg){fig-alt="Line chart of GMT route 6 ridership over time"}

#### Total vehicle revenue hours   

![](export/gmt_vrh_over_time_route_6.svg){fig-alt="Line chart of GMT route 6 vehicle revenue hours over time"}

#### Ridership per vehicle revenue hour 

![](export/gmt_ridership_per_vrh_over_time_route_6.svg){fig-alt="Line chart of GMT route 6 ridership per vehicle revenue hour over time"}

#### Detailed data  

```{r, echo=FALSE}
get_data_table(chrt_dta, "6")
```


---


### 7: North Avenue    

#### Total ridership 

![](export/gmt_ridership_over_time_route_7.svg){fig-alt="Line chart of GMT route 7 ridership over time"}

#### Total vehicle revenue hours

![](export/gmt_vrh_over_time_route_7.svg){fig-alt="Line chart of GMT route 7 vehicle revenue hours over time"}

#### Ridership per vehicle revenue hour

![](export/gmt_ridership_per_vrh_over_time_route_7.svg){fig-alt="Line chart of GMT route 7 ridership per vehicle revenue hour over time"}

#### Detailed data  

```{r, echo=FALSE}
get_data_table(chrt_dta, "7")
```


---  


### 8: City Loop  

#### Total ridership 

![](export/gmt_ridership_over_time_route_8.svg){fig-alt="Line chart of GMT route 8 ridership over time"}


#### Total vehicle revenue hours  

![](export/gmt_vrh_over_time_route_8.svg){fig-alt="Line chart of GMT route 8 vehicle revenue hours over time"}

#### Ridership per vehicle revenue hour 

![](export/gmt_ridership_per_vrh_over_time_route_8.svg){fig-alt="Line chart of GMT route 8 ridership per vehicle revenue hour over time"}

#### Detailed data  

```{r, echo=FALSE}
get_data_table(chrt_dta, "8")
```

---  


### 9: Winooski  

#### Total ridership 

![](export/gmt_ridership_over_time_route_9.svg){fig-alt="Line chart of GMT route 9 ridership over time"}

#### Total vehicle revenue hours 

![](export/gmt_vrh_over_time_route_9.svg){fig-alt="Line chart of GMT route 9 vehicle revenue hours over time"}

#### Ridership per vehicle revenue hour 

![](export/gmt_ridership_per_vrh_over_time_route_9.svg){fig-alt="Line chart of GMT route 9 ridership per vehicle revenue hour over time"}

#### Detailed data  

```{r, echo=FALSE}
get_data_table(chrt_dta, "9")
```


---  


### 11: City Connector  

#### Total ridership 

![](export/gmt_ridership_over_time_route_11.svg){fig-alt="Line chart of GMT route 11 ridership over time"}

#### Total vehicle revenue hours 

![](export/gmt_vrh_over_time_route_11.svg){fig-alt="Line chart of GMT route 11 vehicle revenue hours over time"}

#### Ridership per vehicle revenue hour 

![](export/gmt_ridership_per_vrh_over_time_route_11.svg){fig-alt="Line chart of GMT route 11 ridership per vehicle revenue hour over time"}

#### Detailed data  

```{r, echo=FALSE}
get_data_table(chrt_dta, "11")
```


---  


### 86: Montpelier LINK Express  

#### Total ridership 

![](export/gmt_ridership_over_time_route_86.svg){fig-alt="Line chart of GMT route 86 ridership over time"}

#### Total vehicle revenue hours 

![](export/gmt_vrh_over_time_route_86.svg){fig-alt="Line chart of GMT route 86 vehicle revenue hours over time"}

#### Ridership per vehicle revenue hour 

![](export/gmt_ridership_per_vrh_over_time_route_86.svg){fig-alt="Line chart of GMT route 86 ridership per vehicle revenue hour over time"}

#### Detailed data  

```{r, echo=FALSE}
get_data_table(chrt_dta, "86")
```


---  


### 56 and 96: Franklin County Commuter

Routes 56 (Milton Commuter) and 96 (St. Albans Link) are reported as a single combined route group for consistency, following the absorption of Route 56 into Route 96 to create the Franklin County Commuter on August 25, 2025.

#### Total ridership 

![](export/gmt_ridership_over_time_route_56-96.svg){fig-alt="Line chart of GMT route 56 and 96 ridership over time"}

#### Total vehicle revenue hours 

![](export/gmt_vrh_over_time_route_56-96.svg){fig-alt="Line chart of GMT route 56 and 96 vehicle revenue hours over time"}

#### Ridership per vehicle revenue hour 

![](export/gmt_ridership_per_vrh_over_time_route_56-96.svg){fig-alt="Line chart of GMT route 56 and 96 ridership per vehicle revenue hour over time"}

#### Detailed data  

```{r, echo=FALSE}
get_data_table(chrt_dta, "56-96")
```

<div class="dashboard-footer">
  © 2026 [Green Mountain Transit](https://ridegmt.com/). Data provided as-is and subject to revision.
</div>